
\chapter{Runtime Overheads Migration}
\label{chapter.taskgenx}
As chip multi-processors (CMPs) are becoming more and more complex, software solutions such as parallel programming models are attracting a lot of attention.
Task-based parallel programming models offer an appealing approach to utilize complex CMPs.
However, the increasing number of cores on modern CMPs is pushing research towards the use of fine grained parallelism.
Task-based programming models need to be able to handle such workloads and offer performance and scalability.
Using specialized hardware for boosting performance of task-based programming models is a common practice in the research community.

This chapter makes the observation that task creation becomes a bottleneck when we execute fine grained parallel applications with task-based programming models.
%From our experiments the time spent on per-task generation ranges from 13 to 39 microseconds.
%From our experiments the average time spent on per-task generation is 19,2 microseconds while the average task time is 51936727.
To quantify this statement, the creation of one task lasts from 0,08\% and can reach up to 38,47\% of one task's execution time\footnote{This percentage depends on the task's characteristics such as its data dependencies}.
Applications whose task creation lasts more than 0,5\% of their per-task duration, task creation becomes a bottleneck as there is no continuous provision of tasks to the available resources.
%
%
%each task creation
%From our experiments task duration lasts from 2.5x of task creation up to 29472x of task creation. 
%Applications where their task duration is less than 165x task duration suffer from task creation overheads. 
As the number of cores increases the time spent generating the tasks of the application is becoming more critical to the entire execution.
To overcome this issue, we propose {\proposal}.
{\proposal} offers a solution for minimizing task creation overheads and relies both on the runtime system and a dedicated hardware.
%We propose \proposal, a task-based programming model that offers a minimalistic approach to runtime overheads acceleration.
On the runtime system side, {\proposal} decouples the task creation from the other runtime activities.
It then transfers this part of the runtime to a specialized hardware.
We draw the requirements for this hardware in order to boost execution of highly parallel applications.
From our evaluation using 11 parallel workloads on both symmetric and asymmetric systems, we obtain performance improvements up to 15$\times$, averaging to 3.1$\times$ over the baseline.
\newpage

%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%
\section{Introduction}
\label{sec:intro}
\input{TaskGenX/01-introduction}


%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%
\section{Background and Motivation}
\label{sec.taskgenx.background}
\input{TaskGenX/02-background}

%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%
\section{Task Generation Express (TaskGenX)}
\label{sec.taskgenx.ram}
\input{TaskGenX/03-our_work}

%\section{Hardware requirements}}
%\label{sec:hw}
\input{TaskGenX/hw}
%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%

\input{TaskGenX/04-experimental}

%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%
\section{Evaluation}
\label{sec.taskgenx.evaluation}
\input{TaskGenX/05-evaluation}

%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%
%\section{Related Work}
%\label{sec:related}
%\input{TaskGenX/06-related}

%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%
\section{Conclusions}
\label{sec.taskgenx.conclusions}
\input{TaskGenX/07-conclusions}

