



%M. Casas is supported by the Secretary for Universities and Research of the Ministry of Economy and Knowledge of the Government of Catalonia and the Cofund programme of the Marie Curie Actions of the 7th R\&D Framework Programme of the European Union (Contract 2013 BP\_B 00243).
%M. Moret\'{o} has been partially supported by the Ministry of Economy and Competitiveness under Juan de la Cierva postdoctoral fellowship number JCI-2012-15047.
%M. Moreto has been partially supported by the Ministry of Economy and Competitiveness
%under Ramon y Cajal fellowship number RYC-2016-21104.


\begin{acknowledgements}

First and foremost I would like to express my gratitude to my PhD advisors Dr. Rosa Sala Maria Badia and Dr. Marc Casas Guix. 
This thesis would not be possible without their valuable help and advice throughout this course. 
Their help was often beyond the typical, as they provided me with the encouragement to pursue a PhD and also to believe that we have all it takes to achieve high quality results.

This PhD would not be possible without the help and collaboration of Dr. Alex Rico and Dr. Miquel Moret{\'o}. 
They taught me how to work and be efficient and showed me the way to author great publications by always raising the bar for the quality of our work.
I feel very privileged having worked with such great researchers that always had the answer to every paper rejection and the right next step to be taken.

I want to thank Barcelona Supercomputing Center (BSC) and its director Dr. Mateo Valero for giving me the opportunity to be a part of it.
BSC is a distinguished research center and I feel grateful, not only for the financial support that it offered me during my PhD but also for the opportunity to meet and work closely to accomplished professors and intelligent people.
My friends and colleagues in the Computer Sciences department of BSC (especially the Romol team, the programming models team, and the HCA team) made the workplace and the PhD experience more fun, and I could not miss thanking them for this.

I am very grateful to my friends in Barcelona that are now everywhere.
The bubble-gum gang, Danai, Maria, Michalis, Zhuang, Rajiv, and Rommel as well as Tomasz, Karthik, Pieter, Pade, Linka, Antonia, Tsaliagos, Viola, Maria and Nikos;
they have been by my side during my years in Barcelona and together we had an amazing time, living a second student life.

I could not thank my family enough, Ioanna, David and my sister Deppy that have been supportive during my entire life and filled me with encouragement to continue my PhD studies abroad. 
Roxy, my dog has filled my days with smiles during this period. 
Going home after a paper rejection is hard but taking Roxy to the park after that was always bringing the smile back on my face.
Last but not least, I feel so lucky and thankful for having Giorgos by my side during all my PhD years celebrating my successes, supporting my failures and simply just being always there.




\textit{This work has been supported by the grant SEV-2011-00067 of the Severo Ochoa Program, awarded by the Spanish Government, 
by the Mont-Blanc 2 project (FP7-610402), 
by the RoMoL ERC Advanced Grant (GA 321253), 
by the European HiPEAC Network of Excellence, 
by the Spanish Ministry of Science and Innovation (contracts TIN2012-34557, CAC2007-00052 and TIN2015-65316-P), 
by the Generalitat de Catalunya (contracts 2014-SGR-1051 and 2014-SGR-1272), 
by the European Union's Horizon 2020 research and innovation programme under grant agreement No 671697 and No. 779877 and by the Spanish Government (SEV2015-0493).
The Mont-Blanc project receives funding from the EU’s Seventh Framework Programme (FP7/2007-2013) under grant agreement nº 610402 and from the EU’s H2020 Framework Programme (H2020/2014-2020) under grant agreement nº 671697.}
\end{acknowledgements}